---
output:
  html_document: default
  pdf_document: default
---
```{r}
tgpp<-read.csv("tgpp.csv")
```
1. What are the names of the columns in this dataset?
2. How many rows and columns does this data file have?
```{r}
str(tgpp)
#1 plot, year, record_id, corner, scale, richness, easting, northing, slope, ph, yrsslb
#2 11 columns, 4080 rows
```
3. What kind of object is each data column? Hint: checkout the function sapply().
```{r}
sapply(tgpp, class)
```

4. What are the values of the the datafile for rows 1, 5, and 8 at columns 3, 7, and 10
```{r}
tgpp[c(1,5,8), c(3,7,10)]
```
5. Create a pdf of the relationship between the variables "scale" and "richness". Scale is the area in square meters of the quadrat in which richness was recorded. Be sure to label your axes clearly, and choose a color you find pleasing for the points. To get a list of available stock colors use the function colors().
```{r}
pdf("./scale_richness.pdf")
plot(tgpp$scale, tgpp$richness, xlab = "Scale", ylab="Richness", col="dodgerblue")
```
6. What happens to your plot when you set the plot argument log equal to 'xy'. plot(..., log='xy')
```{r}
plot(tgpp$scale, tgpp$richness, log = "xy", xlab = "Scale", ylab="Richness", col="dodgerblue")
```
 Graph is more spread out and looks nicer. It was log transformed

