---
output:
  html_document: default
  pdf_document: default
---
```{r}
data("iris")
head(iris)
```

```{r}
# create unique vector of species names
sp_ids = unique(iris$Species)

# makes an empty matrix that is 3x4
output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1)
#assigns the species ids as the rownames of the output
rownames(output) = sp_ids
#assigns the variables as the column names
colnames(output) = names(iris[ , -ncol(iris)])

for(i in seq_along(sp_ids)) {
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    for(j in 1:(ncol(iris_sp))) {
        x = 0
        y = 0
        if (nrow(iris_sp) > 0) {
            for(k in 1:nrow(iris_sp)) {
                x = x + iris_sp[k, j]
                y = y + 1
            }
            output[i, j] = x / y 
        }
    }
}
output
```
1.Describe the values stored in the object output. In other words what did the loops create?
  # The loop created the averages for each of the variables for the species.

2.Describe using pseudo-code how output was calculated
  #Loop from 1 to length of species identities
     Take a subset of iris data 
     Loop from 1 to number of columns (traits) of the iris data
        If number of rows is >0 then output then sum f each trait for each species and divide by the          total number of rows for the corresponding species
        
3.The variables in the loop were named so as to be vague. How can the objects output, x, and y could be renamed such that it is clearer what is occurring in the loop
  #output should be species_avg, x should be trait_sum, y should be samp_size

4.It is possible to accomplish the same task using fewer lines of code? Please suggest one other way to calculate output that decreases the number of loops by 1.
```{r}
avg_trait <- matrix(NA, nrow = 3, ncol = 4)
for(i in seq_along(sp_ids)) {
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    for(j in 1:(ncol(iris_sp))) {
      avg_trait[i, j] <- mean(iris_sp[,j])
    }
}
avg_trait
```
5. You have a vector x with the numbers 1:10. Write a for loop that will produce a vector y that contains the sum of x up to that index of x. So for example the elements of x are 1, 2, 3, and so on and the elements of y would be 1, 3, 6, and so on.
```{r}
x<- c(1:10)
y<-NULL
for(i in x){
  y[i]<-sum(x[1:i])
}
y
```

6. Modify your for loop so that if the sum is greater than 10 the value of y is set to NA 
```{r}
x<- c(1:10)
y<-NULL
for(i in x){
  y[i]<-sum(x[1:i])
  if(y[i]>10){
    print("NA")
  }
}
y
```

7. Place your for loop into a function that accepts as its argument any vector of arbitrary length and it will return y.
```{r}
sum_seq <- function(d){
  b<-NULL
  for(i in d){
    b[i]<-sum(d[1:i])
  }
  print(b)
}
sum_seq(x)
```

