```{r}
library(vegan)
data(dune)
data(dune.env)
?dune
```

```{r}
dune_mds <- metaMDS(dune)
```

```{r}
plot(dune_mds, type='n')
text(dune_mds, 'sp', cex=.5)
color_vect = rev(terrain.colors(6))[-1]
points(dune_mds, 'sites', pch=19, 
        col=color_vect[dune.env$Moisture])
legend('topright', paste("Moisture =", 1:5, sep=''), 
        col=color_vect, pch=19)
```

1. It looks like moisture is impacting the species, especially moisture 4 and 5. They are clustered farther away from the other points and moisture 1 and 2 are clustered on the other side of the plot. It would appear that moisture does have some sort of impact on the species.

```{r}
dune.env$Moisture = factor(dune.env$Moisture, levels = 1:5, ordered = FALSE)
cca_dune <- cca(dune ~ ., data=dune.env)
cca_dune1 <- cca(dune ~ dune.env$Moisture + dune.env$A1 + dune.env$Management)
cca_dune2 <- cca(dune ~ dune.env$Moisture + dune.env$A1 + dune.env$Manure)
cca_dune3 <- cca(dune ~ dune.env$Moisture + dune.env$A1 + dune.env$Management + dune.env$Manure)
RsquareAdj(cca_dune, 100)
RsquareAdj(cca_dune1, 100)
RsquareAdj(cca_dune2, 100)
RsquareAdj(cca_dune3, 100)
```

```{r}
anova(cca_dune, permutations = 999)

anova(cca_dune, by = 'margin', permutations = 999)

```

```{r}
plot(cca_dune, type='n', scaling=1)
orditorp(cca_dune, display='sp', cex=0.5, scaling=1, col='blue')
#text(cca_dune, display='bp', col='red')
orditorp(cca_dune, display='cn', col='red')

```

3. From this plot and analysis it complements the mds plot. Moisture is still an important variable. Moisture 5, 1, 3 have clearly distinct clusters. However the cca showed that management and manure also might be an important variable. We didn't see those in the mds, because we only plotted it for moisture. I think the cca analysis is more useful, but the mds plots look nicer and are easier to understand/make, but both are useful analyzes.

```{r}
dune.fit <- envfit(dune_mds, dune.env)
plot(dune_mds, type = 'n')
text(dune_mds, 'sp', cex = .5)
mois_lvs <- sort(unique(dune.env$Moisture))
color_vect <- rev(terrain.colors(length(mois_lvs) + 1))[-1]
points(dune_mds, 'sites', pch=19,
       col = color_vect[dune.env$Moisture])
plot(dune.fit, p.max = 0.01, cex = .75)
legend('topright', paste("Moisture =", mois_lvs, sep = ''),
       col =color_vect, pch=19)
```

