```{r}
library(vegan)
data(BCI)
BCI_xy <- data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))
?BCI
```

Trichilia.tuberculata-common
Trema.micrantha-rare

```{r}
rare <- data.frame(BCI$Trema.micrantha)
com <- data.frame(BCI$Trichilia.tuberculata)
```


```{r}
plot(BCI$Trema.micrantha)
plot(BCI$Trichilia.tuberculata)
```
For Trichilia tuberculata, there doesn't seem to be any sort of spatial dependence. 

For Trema micrantha, I don't see any sort of spatial dependence, but it has a more spotty distribution.


```{r}
c_dist <- dist(com)
r_dist <- dist(rare)
xy_dist <- dist(BCI_xy)
max_dist <- max(xy_dist)/2

plot(xy_dist, c_dist)
abline(lm(c_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, c_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)

plot(xy_dist, r_dist)
abline(lm(r_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, r_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)


c_mantel <- mantel(xy_dist, c_dist)
c_mantel

r_mantel <- mantel(xy_dist, r_dist)
r_mantel
```

From these mantel test, there still doesn't appear to be any type of spatial relationship 



```{r}
sp_a <- c("Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha")
```

```{r}
library(nlme)
dstand <- BCI[ , c('Drypetes.standleyi', sp_a)]
dat <- data.frame(dstand, BCI_xy)
model1 <- gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra +
           Picramnia.latifolia + Quassia.amara +
           Tabernaemontana.arborea + Trattinnickia.aspera + 
           Xylopia.macrantha, data = dat)
model2 <- gls(Drypetes.standleyi ~ Cordia.lasiocalyx , data = dat)
model1
model2
```

```{r}
?Variogram
library(nlme)
plot(Variogram(model1, form = ~ x + y))
plot(Variogram(model2, form = ~ x + y))
```

```{r}
model1_exp = update(model1, corr=corExp(form=~x + y))
model2_exp = update(model2, corr=corExp(form=~x + y))
plot(Variogram(model1_exp, maxDist = max_dist))
plot(Variogram(model1_exp, resType='normalized', maxDist = max_dist))
plot(Variogram(model2_exp, maxDist = max_dist))
plot(Variogram(model2_exp, resType='normalized', maxDist = max_dist))
```


```{r}
mod1_exp_nug = update(model1_exp, corr=corExp(form=~x + y, nugget=T))
mod2_exp_nug = update(model2_exp, corr=corExp(form=~x + y, nugget=T))
plot(Variogram(mod1_exp_nug, maxDist = max_dist))
plot(Variogram(mod1_exp_nug, resType='n', maxDist = max_dist))
plot(Variogram(mod2_exp_nug, maxDist = max_dist))
plot(Variogram(mod2_exp_nug, resType='n', maxDist = max_dist))
```

```{r}
mod1_rat_nug = update(model1, corr=corRatio(form=~x + y, nugget=T))
mod2_rat_nug = update(model2, corr=corRatio(form=~x + y, nugget=T))
plot(Variogram(mod1_rat_nug, maxDist = max_dist))
plot(Variogram(mod1_rat_nug, resType='n', maxDist = max_dist))
plot(Variogram(mod2_rat_nug, maxDist = max_dist))
plot(Variogram(mod2_rat_nug, resType='n', maxDist = max_dist))
```

```{r}
model1
model1_exp
mod1_exp_nug
mod1_rat_nug
```

Adding the error terms affected the coefficients for model 1. The best looking model was the orginial model too, based off the plots.

```{r}
model2
model2_exp
mod2_exp_nug
mod2_rat_nug
```

The error terms changed the coefficients slightly with model 2. Based off the plots, the exponential model with a nugget was the best model 2.

```{r}
anova(model1, model1_exp, mod1_exp_nug, mod1_rat_nug)
anova(model2, model2_exp, mod2_exp_nug, mod2_rat_nug)
```

So the spatial error terms did not really improve the models fit for model 1. It improved the model when compared to the orginal, but the best model was the exponential one. I would assume this is because the model is looking at the effect of multiple species on the species of interest so the error term had less of an effect on the model.

The error term did improve the model for model 2. The best model was the rational quadratic model. The error terms improved model 2 more than they did in model 1. I would assume this is because there was only one species being examined in the model in so error term had more of an effect.

I'm not sure I'm understanding why the error terms influence the models correctly...